\documentclass[conference]{IEEEtran}
\usepackage{cite}

\begin{document}

\title{Indoor Localization for Mobile Devices Using Bluetooth Low Energy Beacons and Wi-Fi Access Points}

\author{\IEEEauthorblockN{Justin L. Sewell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Juan C. Morales}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}}

\maketitle

\begin{abstract}
Due to the weaknesses of GPS signals indoors, a successful method of indoor
navigation is an important research topic. An indoor navigation application
running on a mobile device has numerous uses. The application could help new
students locate their classes or professors and it would especially benefit
disabled students by directing the student in the right direction while taking
the shortest route. This project presents a prototype of an indoor navigation
system using Bluetooth Low Energy beacon devices and mobile phones.
Bluetooth Low Energy (BLE) beacons provide a way of locating a mobile device
within doors. The mobile device has the capability to receive Bluetooth signals
sent out by the beacons. A mathematical model is then used to approximate the
distance between the mobile device and the beacon by using the transmitted
power level and received signal strength of the Bluetooth signal. A grid can
then be used to model a floor of a building with a beacon in each quadrant.
Methods of increasing the accuracy of the application include using a density
algorithm on a sample of received signals and also a probabilistic algorithm
that implements a variation of Bayes theorem.
\end{abstract}

\section{Introduction}
There are more than 7.3 million people in the United States with some form of visual impairment [1]. For many of these individuals, navigation has always been a challenge. The ease of navigation for the blind was greatly improved with the introduction of the Global Positioning System (GPS) and the widespread deployment of systems that use the GPS. However, due to the GPS satellites’ line of sight requirements, these individuals are left unassisted upon entering an indoor environment because no practical system for indoor navigation has been introduced.

The problem with such a system pertains to localizing and tracking the user in an indoor space. This issue has many challenges that must be faced that are highlight in [2], as including: the loss of signal precision of wireless systems due to Non-Line-of-Sight (NLOS) conditions and multipath effect, scaling the system for large spaces, complex environments, and the nonstatic nature of persons and obstacles in indoor settings.

A practical, accurate and cost-efficient indoor navigation system that considers these challenges would greatly benefit the visually impaired. With no form of navigation assistance when in an indoor setting, these individuals are hindered when traversing public spaces, such as malls, universities, airports and bus or train stations, among others. This would mean these individuals will need some form of help to locate his or her desired destination in such structures. An option for such assistance is a navigation application for a mobile device. Such an application would require the user to be localized within an indoor space.

Many methods of indoor localization systems have been explored, using a wide range of hardware based techniques. Previous methods that have been tested use technologies such as GSM (the current global mobile communication standard), radio frequency identification tags (RFID), infrared beacons and receivers, and ultrasonic sensors [3] - [7]. These approaches suffer from various drawbacks such as short detection ranges, high installation costs, unsuitable levels of accuracy, and little space for improvement.

Other more practical solutions to the localization problem use the already in place Wi-Fi infrastructure of buildings to reduce both cost and installation times while maintaining efficient levels of accuracy. Wi-Fi Access points communicate with other devices via 802.11 standards, and the signal strength being received by any device from a Wi-Fi access point, known as the Received Signal Strength Indicator, or RSSI, can be retrieved and used through various techniques to achieve Indoor Localization. [?][Might still have the same issue..]

More recently, approaches that use Bluetooth Low Energy (BLE) beacons have been examined. BLE is a technology that has recently surfaced that is a great candidate for implementing indoor localization due to its low energy consumption, affordability and compact size. Unlike classic Bluetooth, BLE scan times are much shorter, allowing for a quicker refresh rate. In 2015, Google released Eddystone™, an open BLE beacon format that can be configured to send several different types of payloads using the same packet format [8]. [BEFORE EDDYSTONE, THE ONLY FORMAT WAS iBEACON AND ALTBEACON...] The format can be used to create a contextually aware experience for users by delivering proximity event-triggered attachments. Other formats such as iBeacon do not provide this functionality. Eddystone is also more developer friendly and formats can be configured based on context of application.[EDDYSTONE IS MORE DEVELOPER FRIENDLY THAN PAST FORMATS - GIVE FORMAT WEIGHT]

BluNavi localizes the user by fusing data provided by Inertial Measurement Units (IMUs) and distance approximations of the BLE beacons. To further increase the accuracy, the system is complemented by Wi-Fi fingerprinting , a method which makes use of access points by mapping their RSSI values to absolute locations. Eddystone™ configured beacons will be used to drive our mobile, context based, indoor navigation application for the visually impaired. The system will communicate with the user and the beacons/access points through an Android application to provide accurate, real-time indoor navigation. With this approach we aim to provide a low-cost, widely deployable system while still maintaining  a high-level of accuracy.

The rest of this paper is organized as follows: Section II describes current indoor localization research. Section III explains the methodology behind our approach and section IV contains the evaluation of the experimental results. Lastly, Section V details our conclusions and future work.

\section{Related Work}

\section*{Acknowledgment}


The authors would like to thank...


\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}




% that's all folks
\end{document}
